\section{Internal contracts}\label{sec:internal_contract}

Conflux introduces internal contracts for specific usage. A high-level description for the internal contracts is given in Section~\ref{sec:internal}. 
%
Currently, Conflux has three internal contracts with addresses as follows. 
\begin{align}
	a_{\sf admin} \eqdef \admincontract \\ 
	a_{\sf sponsor} \eqdef \sponsorcontract \\
	a_{\sf stake} \eqdef \stakingcontract 
\end{align}
%
When the recipient's address $r$ is one of the internal contracts, Conflux processes $\execute_{\sf internal}(\st^*,g,I)$ and returns $\left(\st^{**}, g^{**},  A, \vec{o} \right)$. 

\subsection{Interfaces and gas required}\label{sec:internal_gas}

In execution of internal contracts, the call data $I_\mathbf{d}$ is interpreted as a function call to Solidity interface. Function $V(\st,c,I_\mathbf{d})$ gives the gas used for internal contract execution by called function and parameters.
$\\$

\begin{tabu}{llll}
	\hline
	Address $I_a$ & Solidity interface & Formal parameters & $V(\st,c,I_\mathbf{d})$ \\
	\hline
	\multirow{2}{*}{$a_{\sf admin}$} & {\tt \small set\char`_admin(address,address)} & $a_0,a_1\in \B_{160}$ & $G_{sset}$\\\cline{2-4}
	& {\tt \small destroy(address)} & $a_0\in \B_{160}$ & $G_{sset}$ \\\hline
	\multirow{4}{*}{$a_{\sf sponsor}$} & {\tt \small set\char`_sponsor\char`_for\char`_gas(address,uint256)}  & $a_0\in \B_{160},n_1\in \N_{256}$ & $G_{sset}$\\\cline{2-4}
	& {\tt \small set\char`_sponsor\char`_for\char`_collateral(address)}  & $a_0\in \B_{160}$ &  $2\times G_{sset}$\\\cline{2-4}
	& {\tt \small add\char`_privilege(address[])} & $\mathbf{a}\in \B_{160}^*$ & $|\mathbf{a}|\times G_{sset}$ \\\cline{2-4}
	& {\tt \small remove\char`_privilege(address[])} & $\mathbf{a}\in \B_{160}^*$ & $|\mathbf{a}|\times G_{sset}$ \\ \hline
	\multirow{3}{*}{$a_{\sf stake}$} & {\tt \small deposit(uint256)}  & $n_0\in \N_{256}$ & $2\times(|\st[s]_{deposit}|+1)\times G_{sset}$\\\cline{2-4}
	& {\tt \small withdraw(uint256)}  & $n_0\in \N_{256}$ &  $2\times|\st[s]_{deposit}|\times G_{sset}$\\\cline{2-4}
	& {\tt \small vote\char`_lock(uint256,uint256)} & $n_0,n_1\in \N_{256}$ & $2\times|1+\tolist(\st[s]_{vote})|\times G_{sset}$ \\\hline
	\multicolumn{4}{l}{where $\st[s]'_{vote}\text{ is defined by }\st[s]_{vote} \text{ removing all the key }x\text{ with }\st[s]_{vote}[x]=\st[s]_{vote}[x-1]$}
\end{tabu}

\subsection{Internal contracts exceptions}
The execution of internal contracts may fail in the following cases 
\begin{itemize}
	\item Conflux parses the call data $I_\mathbf{d}$ as solidity function call. In case the call data doesn't match any solidity function interfaces list as follows, or the format is incorrect, the execution fails. 
	\item The recipient is inconsistent with the code address or internal contract is called by \op{STATICCALL}, i.e., $I_r\neq c$ or $I_w=\bot$. 
	\item The gas $g$ passed in is not enough for internal contract execution, i.e., $g<V(\st^*,c,I_\mathbf{d})$. 
	\item The staking vote contract $a_{\sf stake}$ forbids value transfer to prevent misusing, i.e., $I_a= a_{\sf stake}\;\wedge \; I_v\neq 0$. \footnote{User don't need to transfer balance to staking vote contract in staking. The transferred value to staking vote contract will lost. }
	\item If other exceptions met during the execution, the execution also fails. See the following for details. 
\end{itemize}

Whenever the execution fails, the return values are set as
\begin{align}
	\left(\st^{**}, g^{**},  A, \vec{o} \right)\eqdef (\varnothing,0,A^0,\varnothing)
\end{align} 

If the execution is successful, the remained gas and return data are set as 
\begin{align}
	\left(g^{**}, \vec{o} \right)\eqdef (g-V(\st^*,c,I_\mathbf{d}),\emptystring)
\end{align}

The resultant world-state $\st^{**}$ and substate $A$ is computed depending on interfaces. 


\subsection{Admin Contract}

The admin contract with address $a_{\sf admin}$ gives two interfaces.

\subsubsection{Set administration}

For interface {\tt set\char`_admin(address,address)}, let $a_0,a_1\in \B_{160}$ be the address parameters. 
	\begin{align}
		A & \eqdef A^0 \\
		\st^{**} &\eqdef \st^* \qquad \mbox{except:}\\
		\st^{**}[a_0]_a &\eqdef
			a_1 \qquad \text{if } \mathsf{Type}_{a_0} = [1000]_2 \;\wedge\; I_o = \st^*[a_0]_a 
    \end{align}
    
\subsubsection{Destory contract} 

For interface {\tt destroy(address)}, let $a_0\in \B_{160}$ be the address parameters. The execution fails if 
	\begin{align}
			\st^*[a_0]_o>0\;\wedge\; 	I_o = \st^*[a_0]_a
	\end{align}
	If the execution not fails, 
	\begin{align}
		(\st^{**},A) \eqdef \left\{\begin{array}{ll}
			(\st^*,A^0) & I_o \neq \st^*[a_0]_a \\
			\Psi(\st^*,A^0) & I_o = \st^*[a_0]_a
		\end{array}\right.
	\end{align}
	where $\Psi$ is defined in section~\ref{sec:contract_destruct}.

\subsection{Sponsorship Contract}
The sponsorship contract with address $a_{\sf sponsor}$ gives four interfaces.

\subsubsection{Set sponsor for gas}

For interface {\tt set\char`_sponsor\char`_for\char`_gas(address,uint256)}, let $a_0\in \B_{160},n_1\in \N_{256}$ be the address parameters. The execution fails if 
	\begin{align}
		& \st^*[a_0]=\emptyset \;\vee\; \mathsf{Type}_{a_0}\neq[1000]_2 \;\vee\; \st^*[a_{\sf sponsor}]_b < 1000\times n_1 \;\vee\; n_1\le \st^*[a_0]_p[{\sf limit}]\le \st^*[a_0]_p[{\sf gas}]_b \notag \\ 
		& \;\vee\; (\st^*[a_0]_p[{\sf gas}]_a \neq I_s \;\wedge\; \st^*[a_{\sf sponsor}]_b\le \st^*[a_0]_p[{\sf gas}]_b)  
	\end{align}
	If the execution not fails, 
		\begin{align}
			A & \eqdef A^0 \\
			\st^{**} &\eqdef \st^* \qquad \mbox{except:}\\
			\st^{**}[a_{\sf sponsor}]_b &\eqdef 0 \\ 
			\st^{**}[a_0]_p[{\sf limit}] &\eqdef n_1\\ 
			\st^{**}[a_0]_p[{\sf gas}]_a &\eqdef a_0\\ 
			\st^{**}[a_0]_p[{\sf gas}]_b &\eqdef \begin{cases}
				\st^*[a_{\sf sponsor}]_b + \st^*[a_0]_p[{\sf gas}]_b& p = I_s\\ 
				\st^*[a_{\sf sponsor}]_b& p \neq I_s
			\end{cases}\\ 
			\st^{**}[p]_b &\eqdef \begin{cases}
				\st^*[p]_b & p = I_s\\ 
				\st^*[p]_b + \st^*[a_0]_p[{\sf gas}]_b& p \neq I_s
			\end{cases}\\
			\mbox{where:}& \\
			p & \eqdef \st^*[a_0]_p[{\sf gas}]_a
		\end{align}
    
\subsubsection{Set sponsor for collateral}

For interface {\tt set\char`_sponsor\char`_for\char`_collateral(address)}, let $a_0\in \B_{160}$ be the address parameter. The execution fails if 
	\begin{align}
		& \st^*[a_0]=\emptyset \;\vee\; \mathsf{Type}_{a_0}\neq[1000]_2 \;\vee\; \st^*[a_{\sf sponsor}]_b = 0  \notag \\ 
		& \;\vee\; (\st^*[a_0]_p[{\sf col}]_a \neq I_s \;\wedge\; \st^*[a_{\sf sponsor}]_b\le \st^*[a_0]_p[{\sf col}]_b)  
	\end{align}
	If the execution not fails, 
		\begin{align}
			A & \eqdef A^0 \\
			\st^{**} &\eqdef \st^* \qquad \mbox{except:}\\
			\st^{**}[a_{\sf sponsor}]_b &\eqdef 0 \\ 
			\st^{**}[a_0]_p[{\sf col}]_a &\eqdef a_0\\ 
			\st^{**}[a_0]_p[{\sf col}]_b &\eqdef \begin{cases}
				\st^*[a_{\sf sponsor}]_b + \st^*[a_0]_p[{\sf col}]_b& p = I_s\\ 
				\st^*[a_{\sf sponsor}]_b& p \neq I_s
			\end{cases}\\ 
			\st^{**}[p]_b &\eqdef \begin{cases}
				\st^*[p]_b & p = I_s\\ 
				\st^*[p]_b + \st^*[a_0]_p[{\sf col}]_b& p \neq I_s
			\end{cases}\\
			\mbox{where:}& \\
			p & \eqdef \st^*[a_0]_p[{\sf col}]_a
        \end{align}
        
\subsubsection{Add addresses to whitelist}

For interface {\tt add\char`_privilege(address[])}, let $\mathbf{a}\in \B_{160}^*$ be the address list parameter. The execution fails if \begin{align}
		\mathsf{Type}_{I_s}\neq[1000]_2. 
	\end{align}
	If the execution not fails,
	\begin{align}
		\left(\st^{**}, A\right) &\eqdef \left(\st^{(n)}, A^0\right)\\ 
		\mbox{where:} &\\
		n&\eqdef |\mathbf{a}| \\
		\st^{(0)} &\eqdef \st^* \\ 
		\forall j \in [n],\; \st^{(j)} &\eqdef \Phi(\st^{(j-1)},a_{\sf sponsor},I_s\cdot {\bf a}[j-1],1,I_i)
	\end{align}
	Function $\Phi$ is defined in Section~\ref{sec:storage_maintain}.
    
\subsubsection{Remove addresses to whitelist}

For interface {\tt remove\char`_privilege(address[])}, let $\mathbf{a}\in \B_{160}^*$ be the address list parameter. The execution fails if
 \begin{align}
		\mathsf{Type}_{I_s}\neq[1000]_2. 
	\end{align}
	If the execution not fails, 
	\begin{align}
		\left(\st^{**}, A\right) &\eqdef \left(\st^{(n)}, A^0\right)\\ 
		\mbox{where:} &\\
		n&\eqdef |\mathbf{a}| \\
		\st^{(0)} &\eqdef \st^* \\ 
		\forall j \in [n],\; \st^{(j)} &\eqdef \Phi(\st^{(j-1)},a_{\sf sponsor},I_s\cdot {\bf a}[j-1],0,I_i)
    \end{align}
    
\subsection{Staking vote contract}
The staking vote contract with address $a_{\sf stake}$ gives three interfaces:

\subsubsection{Staking}

For interface {\tt deposit(uint256)}, let $n_0\in \N_{256}$ be the integer parameter. The execution fails if 
	\begin{align}
		n_0 < 10^{18} \;\vee\; \st^*[I_s]_b<n_0
	\end{align}
	If the execution not fails, 
	\begin{align}
		A & \eqdef A^0 \\
		\st^{**} &\eqdef \st^* \qquad \mbox{except:}\\
		\st^{**}[I_s]_b &\eqdef \st^*[I_s]_b - n_0\\ 
		\st^{**}[I_s]_t &\eqdef \st^*[I_s]_t + n_0\\ 
		\st^{**}[I_s]_{\bf d} &\eqdef \st^*[I_s]_{\bf d} \cdot (n_0,|I_\mathbf{L}|,\st^*[a_{\sf sponsor}]_{\bf s}[k_1]_v) \\
		\st^{**}[a_{\sf sponsor}]_{\bf s}[k_2]_v & \eqdef\st^*[a_{\sf sponsor}]_{\bf s}[k_2]_v + n_0 \\
		\mbox{where:}& \\
		k_1 & \eqdef [{\sf accumulate\char`_interest\char`_rate}]_{\sf ch} \\
		k_2 & \eqdef [{\sf total\char`_staking\char`_tokens}]_{\sf ch}
    \end{align}
    
\subsubsection{Lock staking to obtain vote power}

For interface {\tt vote\char`_lock(uint256,uint256)}, let $n_0,n_1\in \N_{256}$ be the integer parameters. The execution fails if 
	\begin{align}
		n_1 \le  |I_\mathbf{L}| \;\vee\; \st^*[I_s]_t<n_0
	\end{align}
	If the execution not fails, 
	\begin{align}
		A & \eqdef A^0 \\
		\st^{**} &\eqdef \st^* \qquad \mbox{except:}\\
		\forall x\le |I_\mathbf{L}|, \st^{**}[I_s]_{\bf v}[x] &\eqdef \st^*[I_s]_{\bf v}[|I_\mathbf{L}|+1]\\ 
		\forall |I_\mathbf{L}|<x\le n_1, \st^{**}[I_s]_{\bf v}[x] &\eqdef \max\{\st^*[I_s]_{\bf v}[x],n_0\}
    \end{align}
    
\subsubsection{Withdraw}

For interface {\tt withdraw(uint256)}, let $n_0 \in \N_{256}$ be the integer parameter. The execution fails if 
	\begin{align}
		n_0>\st[I_s]_t - \st[I_s]_{\bf v}[|I_\mathbf{L}|+1]
	\end{align}
	If the execution not fails, 
	\begin{align}
		A & \eqdef A^0 \\
		\st^1 &\eqdef \st^* \qquad \mbox{except:}\\
		\forall x\le |I_\mathbf{L}|, \st^1[I_s]_{\bf v}[x] &\eqdef \st^*[I_s]_{\bf v}[|I_\mathbf{L}|+1]\\ 
		\st^1[I_s]_t &\eqdef \st^*[I_s]_t - n_0\\ 
		\forall i < |\st^*[I_s]_{\bf d}|, \st^1[I_s]_{\bf d}[i][{\sf amt}] &\eqdef \st^*[I_s]_{\bf d}[i][{\sf amt}] - z_i \\
		\st^1[I_s]_b &\eqdef \st^*[I_s]_b + n_0 + q\\
		\st^1[I_s]_r &\eqdef \st^*[I_s]_r + q\\
		\st^1[a_{\sf sponsor}]_{\bf s}[k_2]_v & \eqdef\st^*[a_{\sf sponsor}]_{\bf s}[k_2]_v - n_0 \\
		\st^1[a_{\sf sponsor}]_{\bf s}[k_3]_v & \eqdef\st^*[a_{\sf sponsor}]_{\bf s}[k_3]_v + q \\
		\st^{**} &\eqdef \st^1 \qquad \mbox{except:}\\
		\st^{**}[I_s]_{\bf d} &\eqdef \mbox{Remove elements $e$ from $\st^1[I_s]_{\bf d}$ with $e[{\sf amt}]=0$} \\
		\mbox{where:}& \\
		y_i & \eqdef \sum_{j=0}^{i-1} \st^*[I_s]_{\bf d}[j][{\sf amt}] \\
		z_i & \eqdef \max\{0,\min\{n_0-y_i,\st^*[I_s]_{\bf d}[i][{\sf amt}]\}\} \\
		q & \eqdef \sum_{i=0}^{|\st^*[I_s]_{\bf d}|-1} \left\lfloor \frac{z_i \times \st^*[a_{\sf sponsor}]_{\bf s}[k_1]}{\st^*[I_s]_{\bf d}[i][{\sf accIR}]}\right\rfloor \\
		k_1 & \eqdef [{\sf accumulate\char`_interest\char`_rate}]_{\sf ch} \\
		k_2 & \eqdef [{\sf total\char`_staking\char`_tokens}]_{\sf ch} \\ 
		k_3 & \eqdef [{\sf total\char`_issued\char`_tokens}]_{\sf ch}
	\end{align}


